---
title: "Discussion 6"
author: "Jaideep Bollu"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

For this assignment, you will be using the `mtcars` dataset to answer the questions below that are related to statistics in R, regression, and classification. Functions and other topics will also be included below.

1) Create a function named `statistics` that takes in a single numeric vector, representing a variable/column from the dataset, and returns a single named list containing the following:
- `mean`: the mean of the variable
- `median`: the median of the variable
- `q90`: the 90th percentile of the variable
- `n`: the number of elements in the variable (1.5 pts)

```{r}
statistics <- function(x) {
  xe <- list(mean(x), median(x), quantile(x, probs = 0.9), n = length(x))
  names(xe) <- c("mean", "median", "g90", "n")
  return(xe)
}
```

2) Use the `statistics` function created in Q1 on any two continuous variables of your choice from `mtcars`. Output those results. (1 pt)

```{r}
data("mtcars")
statistics_mpg <- statistics(mtcars$mpg)
statistics_hp <- statistics(mtcars$hp)

statistics_mpg
statistics_hp
```

3) With `hp` as the explanatory variable and `mpg` as the response variable, create a linear regression model and show a summary of the model fit. (1 pt)

```{r}
lrg <- lm(mpg ~ hp, data = mtcars)
lrg
```

4) Suppose you have the following dataset with these `hp`, `mpg`, and `am` values:

```{r}
newMPG = c(26, 29, 13, 18)
newHP = c(150, 125, 250, 160)
newAM = c(0, 0, 1, 1)
predData = data.frame(mpg = newMPG, hp = newHP, am = newAM)
```

Generate `mpg` predictions for the `predData` dataset using the model created in Q3 and show these predictions. Find the mean squared error. (1.5 pts)

```{r}
prediction <- predict(lrg, newdata = predData)
mse <- mean((prediction - predData$mpg)^2)
mse
```

5) With `hp` as the explanatory variable and `am` as the response variable, create a logistic regression model and show a summary of the model fit. (1 pt)

```{r}
log_rm <- glm(am ~ hp, data = mtcars, family = binomial)
summary((log_rm))
```

6) Using `predData`, generate `am` classifications using the model created in Q5 and show these classifications. Find the prediction accuracy. (1.5 pts)

```{r}
pred_probability <- predict(log_rm, newdata = predData, type = "response")
pred_pos <- ifelse(pred_probability < 0.5, 0, 1)
predict_accuracy <- mean(pred_pos == predData$am)
predict_accuracy
```
