---
title: "Discussion 3"
author: "Jaideep Bollu"
output:
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Use built-in *rock* data set in R to answer the questions on this discussion. This data set contains measurements on 48 rock samples from a petroleum reservoir (area = area of pores, peri = total perimeter of pores, shape = perimeter/sqrt(area), perm = permeability).

1) Store the *shape* and *peri* columns of the *rock* data set as vectors called `shape_vec` and `peri_vec` respectively. (1 pt)

```{r}
data("rock")
shape_vec <- rock$shape
peri_vec <- rock$peri
```

2) Look at the four indexing techniques for atomic vectors using the `[]` operator in Section 5.1 of the Github notes. Use two of these techniques to isolate and display all fares in `peri_vec` that are larger than (arithmetic) mean of `peri_vec`. That is, create the same output two different ways. (2 pts)

```{r}
perimeter_mean <- mean(peri_vec)
peri_vec[peri_vec > perimeter_mean]

index <- which(peri_vec <= perimeter_mean)
peri_vec[-index]
```

3) Look at the three operators `[]`,  `[[]]`, `$` in Section 5.4 of the Github notes.  Use all three operators to isolate and display 24th value in the `area` column. That is, create the same output at least three different ways, one for each operator. (3 pts) 

```{r}
val_24 <- rock[24, "area"]
val_24

val_24_2 <- rock[["area"]][24]
val_24_2

val_24_3 <- rock$area[24]
val_24_3
```

4) Using the subsetting method of your choice, subset the *rock* data set so that you only have observations where the *shape* is less than 0.2 and the *perm* is equal to 100. (1.5 pts)

```{r}
rock_subset <- subset(rock, shape < 0.2 & perm == 100)
rock_subset
```
