---
title: "Homework #1 (40 pts)"
author: "Jaideep Bollu"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
 
## Instructions
- If you wish to use any additional packages beyond "Base R" (the default) then email Jericho (`jlaws011@ucr.edu`) for permission. 
- Use R to answer the questions below. 
- Check Canvas regularly for clarification on questions; there may be important posts that will not be announced through *Elearn*.  

## Background Information 

Descriptive statistics describe or summarize attributes of a data set. The one of the most common attributes of interest are central tendency and dispersion. Central tendency refers to the point around which data values tend to accumulate. Dispersion refers to how spread out or disperse the data appear around the accumulation point. There are several **measures of central tendency** and **measures of dispersion**.

### Measure of Central Tendency 
A measure of central tendency is a statistic that describe the middle most value of a data set, or a typical value that you would expect.  Most people think of calculating the (arithmetic) mean when they want to determine the average value.  However, there are several types of averages that can be calculated.  These "averages" are used in different settings for different reasons. We will go over: arithmetic mean, geometric mean, harmonic mean, median 
    
#### Arithmetic Mean
Usually referred to as simply the **mean**, the **arithmetic mean** or **sample mean** is denoted by $\overline{x}$ and is an empirical value computed from the data.  Let $x_1, ..., x_n$ represent the randomly sampled data values, where $n$ is the **sample size** (number of observations).  The arithmetic mean is computed using 
   
  $$\overline{x} = \frac{\sum_{i=1}^n x_i}{n}$$
    
**EXAMPLE**: The total length of four fish fry were 5, 7, 10, 6 mm. Here $n=4$
   
   $$\overline{x} = \frac{5+7+10+6}{4} = \frac{28}{4} = 7$$

#### Geometric Mean
The geometric mean is sometimes used for data that measure proportional changes.  We will let $G$ denote the geometric mean. Let $x_1, ..., x_n$ represent the data values and assume all $x_i>0$.  The geometric mean is computed using 
    
$$G = \exp \left (\frac{\sum_{i=1}^n \ln x_i}{n} \right ) $$
    
where $\ln(x)$ is the natural logarithm $\log_e$ and $\exp(x)$ is the exponential function $(e^x)$. 

**EXAMPLE**:  The average number of offspring per rabbit over three consecutive years was 4.5, 2.0, 3.0 offspring/rabbit.  Here $n=3$.  The geometric mean number of offspring per year is 
    
$$G = \exp \left( \frac{\ln(4.5) + \ln(2.0) + \ln(3.0)}{3} \right) $$
    
$$ = \exp \left( \frac{1.504077+ 0.693147 +1.098612}{3} \right) $$
    
$$= \exp \left( \frac{3.295836}{3} \right)$$
    
$$= \exp(1.09861) $$
    
approximately 3.00 offspring per year 

#### Harmonic Mean 

The **harmonic mean** is sometimes used for data that involve ratios or rates. We will let $H$ denote the harmonic mean. Let $x_1, ..., x_n$ represent the data values and assume all $x_i>0$.  The harmonic mean is computed using 
    
$$H = \left ( \sum_{i=1}^n x_i^{-1} /n\right )^{-1} $$
    
**EXAMPLE**: The doubling times for three strains of bacteria are 1.5, 1.0, 3.0, hours.  Here $n=3$.  The harmonic mean of the doubling times is 
    
$$H = \left ( \frac{(1.5)^{-1} + (1.0)^{-1} + (3.0)^{-1}}{3} \right )^{-1} $$
    
$$ = \left ( \frac{0.666667+ 1.000000 + 0.333333}{3}\right )^{-1} $$
        
$$ = \left ( \frac{2.000000}{3}\right )^{-1} =1.5 \text{hours} $$
    
In this example, $H$ is the doubling time of an equal mixture of the three strains.  The arithmetic mean would overestimate the doubling time ($\overline{x}=1.83$). Note that it is always true that $\overline{x}\geq G \geq H$.  The equality only holds when all the data values are identical. If two or more data values differ, $\overline{x}> G > H$.


#### Median 

The **median** is the middle value in a set of observations.  The computation of the median depends on whether the number of observations is even or odd.  We will let $M$ represent the median. First one must rank the observations.  Assume $x_1\leq x_2 \leq ... \leq x_n$ are the data where each subscript is the rank.  The median is computed using 
    
$$M = \left\{\begin{matrix}
          x_{(n+1)/2} & \text{, if }n\text{ is odd} \\ 
          \frac{x_{n/2} +x_{1+(n/2)}}{2} & \text{, if }n\text{ is even} 
          \end{matrix}\right.$$
            
In other words, if $n$ is odd, $M$ is the middle of the ranked observations, whereas if $n$ is even, $M$ is the average of the two middle observations. 
  
**EXAMPLE**:  The weights of seeds are 2, 11, 1, 5 mg.  Here $n=4$.  After ranking the data we have $x_1 = 1, x_2 = 2, x_3 = 5, x_4 =11$.  Since $n$ is even 

$$M = \frac{x_{4/2} +x_{1+(4/2)}}{2}$$
$$ = \frac{x_{2} +x_{3}}{2}$$
$$ = \frac{2+5}{2}   = 3.5 mg$$
  
One advantage of the median is that it is not sensitive to **outliers** which are extremely large or small data values.  The arithmetic mean can be unduly influenced by extreme outliers, especially if $n$ is small. 

### Measures of Dispersion

A measure of dispersion, sometimes also called a measure of spread, is used to describe the variability of a data set. We usually report a measure of dispersion with a measure of central tendency.  A measure of central tendency (usually the mean) tells us what a typical value is the data set, and a measure of dispersion indicates by how much values tend to deviate from this typical value. 
  
        
#### Variance

The **variance** is one of the most common measures of dispersion.  It does not have a particularly intuitive interpretation, but has ideal theoretical properties. Let $x_1, ..., x_n$ represent the data.  The variance is computed using either of the following formulas
    
$$s^2 = \frac{\sum_{i=1}^n (x_i - \overline{x})^2}{n-1} = \frac{\sum_{i=1}^n x_i^2 - \left ( \left ( \sum_{i=}^n x_i\right )^2/n\right ) }{n-1} $$
Note that these two formulas for variance are identical mathematically.  

**EXAMPLE** : The total length of four fish fry were 5,7, 10, 6 mm.  Here $n=4$. 
    
$$s^2 = \frac{(5)^2 + (7)^2 + (10)^2 + (6)^2 - \left( \left ( 5+7+10+6\right )^2/4 \right ) }{4-1} $$
    
$$ = \frac{25+ 49+ 100+36 - \left ( (28)^2/4\right )}{3}$$
    
$$ = \frac{210 - (784/4)}{3} = \frac{14}{3} = 4.67 mm$$



#### Standard Deviation

The **standard deviation** is also very common, it is the square root of the variance. Let $x_1, ..., x_n$ represent the data.  
    

**EXAMPLE** : The total length of four fish fry were 5,7, 10, 6 mm.  Here $n=4$. We saw above that $s^2 = 4.67$. Thus the standard deviation is $s = \sqrt{4.67} = 2.16$.



#### Coefficient of variation

The **coefficient of variation** is a dimensionless quantity.  It measures dispersion relative to the mean.  We will denote the coefficient of variation as $CV$. Let $x_1, ..., x_n$ represent the data values and assume all $x_i>0$.  The coefficient of variation is computed as the square root of the variance divided by the mean. 
    
$$CV = \frac{s}{\overline{x}}$$
    
  
**EXAMPLE**: The total length of four fish fry were 5, 7, 10, 6 mm.  Here $n=4$. The values for  $\overline{x}=7$ and $s = 2.166667$, calculated above. Therefore, 
    
$$CV = \frac{2.166667}{7.0} = 0.309 $$
        
or $30.9\%$
    
Because it is a dimensionless quantity, the coefficient of variation can be used for making comparisons of dispersion for measurements of different scales; for example, the relative variation in body mass for mice versus elephants or the relative variation in height versus weight among adults. 


#### Range

The **range** is the difference between the largest and the smallest observations.  We will denote the range as $R$. Let $x_1, ..., x_n$ represent the data values.  The range is computed using
    
$$R = \max(x_i) - \min(x_i)$$
    
where $\max()$ and $\min()$ represent the largest and smallest of the observations.      

**EXAMPLE**: The total length of four fish fry were 5, 6, 10, 6 mm.  The range is 
    
$$R = \max(5, 7, 10, 6) - \min(5, 7, 10, 6)$$
    
$$ = 10 - 5 = 5mm$$
    
Because it involves the extreme observations, the range is very sensitive to outliers. 


#### Mean Deviaiton 

The **mean deviation** is the average of the absolute values of the differences between the data and the mean.  We will denote the mean deviation as $MD$. Let $x_1, ..., x_n$ represent the data values.  The mean deviation is computed using 
    
$$MD = \frac{\sum_{i=1}^n |x_i - \overline{x}|}{n}$$
    
where the notation $|x|$ means the absolute value of x. 

**EXAMPLE**: The total length of four fish fry were $5, 7, 10, 6$ mm.  Here $n=4$.  In section 2.2.1 it was shown that $\overline{x} = 7.0$.  The mean deviation is 
    
$$MD = \frac{|5-7| + |7-7| + |10-7| + |6-7|}{4}$$
    
$$=\frac{2+0+3+1}{4} = \frac{6}{4} = 1.5 \text{mm}$$
    
Use of the mean deviation is uncommon.  It appears more often in physics and chemistry. 

## Questions 

**Q1: R Basics**

Suppose we want to assign some variable to be the number 5, but we do not know what to name the variable. Run through the following lines of code below to see which object names are acceptable:

```{r, eval = F}
T = 5
1first = 5
arg = 5
_start = 5
a = 5
```

(a) Based on the initial code given, state which object names we should not use. For each name that failed, explain why these names are not valid.

We should not use the object name 1first because object names can start with a letter or a period and in this instance this object name starts with a digit which is not valid.

(b) After running the following line of code below, determine what the `%%` operator does based on context or through some source (please list the source). :
```{r}
12 %% 5
```

The %% operator finds the modulo operation of the following numbers which we can see in this context as 12 modulo 5 returns a value of 2 which is what is returned by the 12 %% 5 line of code.

(c) Suppose you assign `NUM` to equal some number. Write a line of code using `NUM` and the `%%` operator to determine if a number is even or odd. On the following line of code, type `NUM` to output your result. Explain what the output represents.

```{r}
NUM = 3
NUM <- NUM %% 2
NUM
```

We can determine whether the NUM object is an even or odd number using the NUM %% 2 line of code as it will return the remainder of the value after being divided by 2 and if the line of code returns 0 we find that the NUM value is divisible by 2 and thus even while if the line of code returns 1 we find that it is not divisible by 2 and thus an odd number.

(d) Suppose you need to calculate your final grade in STAT 107. Write a single line of code to calculate your course grade using operators and assign the calculation to a variable of your choice. Type the name of the variable so that it is displayed.  Use the following grades in each group (with its weight):
+ Grades:
  - Discussion (15\%): 0.876
  - Group Activities and Quizzes (15\%): 0.805
  - Homework (40\%): 0.943
  - Final (30\%): 0.823 
  
```{r}
final_grade <- (0.15 * 0.876) + (0.15 * 0.805) + (0.4 * 0.943) + (0.3 * 0.823)
final_grade
```

(e) Using part (d), write a line of code for each condition below:
+ Conditions:
  - Your course grade is greater than 90\%
  - Your course grade is between 80\% and 90\% (inclusive)
  - Your course grade is exactly equal to 85\%
  
```{r}
final_grade > 0.9
final_grade <= 0.9 && final_grade >= 0.8
final_grade == 0.85
```

**Q2: Fibonacci Sequence (Vectors)**

The Fibonacci sequence appears in many biological structures. It is formed by starting with the number 0, 1 and then adding the last two numbers to get the next number: 0, 1, 1, 2, 3, 5, 8, etc.

(a) Create an *integer* vector called `int_fib` that contains the first 10 values of the Fibonacci sequence. Type the name of the vector on the following line so that it is displayed. 

```{r}
int_fib <- c(0L, 1L, 1L, 2L, 3L, 5L, 8L, 13L, 21L, 34L)
int_fib
```

(b) Similarly to part (a), create a *double* type vector called `dbl_fib` with the first 5 values of the Fibonacci sequence.  

```{r}
dbl_fib <- c(0, 1, 1, 2, 3)
dbl_fib
```

(c) Create a *character* vector with any elements of your choice that has length 5. 

```{r}
char_vector <- c("Mango", "Ruler", "Google", "Computer", "California")
char_vector
```

(d) Convert and redefine `int_fib` into a character vector.

```{r}
int_fib2 <- as.character(int_fib)
int_fib2
```

(e) Use a built in function to determine how many elements are in `int_fib`, `dbl_fib`, and your vector in part (c).

```{r}
length(int_fib)
length(dbl_fib)
length(char_vector)
```

(f)  Output the second element of `int_fib` using its index. 

```{r}
int_fib[2]
```

**Q3: Coercion and Vector Types**

Vectors built using `c()` must be all of the same type.  If given a combination of different types of elements, R will convert all the elements to be of the same type using a series of internal rules. 

(a) Create a series of vectors that are composed of elements of different types and use the `typeof()` function to see how R decides what type of vector to create.  You do not need to show intermediate steps or R code here, just fill out your results after the "-->". The first one is filled out for you. (3 points)
+ Combinations:
  - integer, double  --> double 
  - integer, character  --> character
  - integer, logical --> integer
  - double, character --> character
  - double, logical --> double
  - character, logical --> character
  - integer, double, character --> character
  - integer, double, logical --> double
  - integer, character, logical --> character
  - double, character, logical --> character
  - double, logical, integer, character --> character
  

(b) Based on your results in part (a), determine the hierarchy for how R prioritizes different data types.

The hierarchy would be character > double > integer > logical

**Q4: Factors with `mtcars`**

Use the data frame `mtcars`, which is a built in data set in R, to answer the questions below.  Display the final objects created/adjusted for each question. 

```{r, eval = F}
mtcars
```

(a) Create a factor vector called `vs` from the `vs` column of `mtcars`. Make sure it has two levels: 0, 1. 

```{r}
vs <- factor(mtcars$vs, levels = c(0, 1))
vs
```

(b) Create a factor vector called `carb` from the `carb` column of `mtcars`.Have this be an ordered factor with levels:  1 < 2 < 3 < 4 < 5 < 6 < 7 < 8. HINT: Some of these levels do not appear automatically. 

```{r}
carb <- factor(mtcars$carb, levels = c(1:8), ordered = TRUE)
carb
```

(c) Create a frequency table for the column `mpg`.

```{r}
mpg_table <- table(mtcars$mpg)
```

(d) Create a two-way frequency table for the factor vectors `vs` and `carb`. 

```{r}
two_way_table <- table(vs, carb)
two_way_table
```

(e) Redefine the vectors `vs` and `carb` to be character vectors. 

```{r}
vs2 <- as.character(vs)
carb2 <- as.character(carb)
```

(f) Create a two way frequency table for `vs` and `carb` again.

```{r}
char_two_way_table <- table(vs2, carb2)
char_two_way_table
```

(g) Comment on the difference between the frequency table in parts (d) and (f). 

One difference I saw between the two frequency tables would be that the two way frequency table with the vectors redefined as character vectors was missing the values of 5 and 7 for the carb headers of the table which both had values of only zero for the vs catagories of 0 and 1 thus representing that there were no values of 5 and 7 in the carb factor vector.

(h) Display everything inside the vector `carb` except the second to last element. 

```{r}
carb[-(31)]

```


**Q5: Measures of Center**  

For these questions use only (arithmetic/logical/relational/assignment) operators, and the functions `exp()`, `sum()`, `log()`, and `length()`. Do not use `mean()`. Use the vector `LH` below in R to answer the following questions.  This vector contains annual measures of the level (in feet) of Lake Huron, USA.

```{r}
LH <- c(LakeHuron)
```

(a) Calculate the arithmetic mean of the vector `LH`. 

```{r}
arithmetic_mean <- sum(LH) / length(LH)
```

(b) Calculate the geometric mean of vector `LH`.

```{r}
geometric_mean <- exp(sum(log(LH)) / length(LH))

```

(c) Calculate the harmonic mean of the vector `LH`. 

```{r}
harmonic_mean <- (sum(1 / LH) / length(LH))^(-1)
```

(d) Store the geometric mean, harmonic mean, and arithmetic mean of the vector `LH` in a vector called `LH_averages`. 

```{r}
LH_averages <- c(arithmetic_mean, geometric_mean, harmonic_mean)
```

(e) Name each element in the vector `LH_averages` to clearly indicate which statistic is which.

```{r}
names(LH_averages) <- c("arithmetic mean", "geometric mean", "harmonic mean")
```

(f) Use a one line command to check that arithmetic mean is larger than the geometric mean AND that the geometric mean is larger than the harmonic mean. 

```{r}
arithmetic_mean > geometric_mean & geometric_mean > harmonic_mean
```

(g) Use a one-line command to find the minimum of the vector `LH_averages`.

```{r}
min(LH_averages)
```

(h) The `which` function can be used to identify which indices the inputted Boolean statement occurs at. Use the `which` function along with the command from part (g) to find which index the minimum occurs at in `LH_averages`.  
```{r}
which(LH_averages == min(LH_averages))
```

(i) Calculate the (arithmetic) mean, median, geometric mean, and harmonic mean for the vector `cyl` from the `mtcars` dataset. (You can now use any base R function for this.)  
`
```{r}
arithmetic_mean_cyl <- mean(mtcars$cyl)
geometric_mean_cyl <- exp(sum(log(mtcars$cyl) / length(mtcars$cyl)))
harmonic_mean_cyl <- (sum(1 / mtcars$cyl) / length(mtcars$cyl))^(-1)
median_cyl <- median(mtcars$cyl)
```

(j) Store the results of the (arithmetic) mean, median, geometric mean, and harmonic mean for `cyl` in a *named* list called `measures_of_center`. The names should indicate which statistic is which. 

```{r}
measures_of_center <- list(
  Arithmetic_Mean = arithmetic_mean_cyl,
  Median = median_cyl,
  Geotmetric_Mean = geometric_mean_cyl,
  Harmonic_Mean = harmonic_mean_cyl
)
```

**Q6: Measures of Spread**

For these questions use only (arithmetic/logical/relational/assignment) operators, and the functions `exp()`, `sum()`, `log()`, `abs()`, `length()`. Do not use `mean()` or `var()`. Use the vector `LH` from above to answer the following questions.

(a) Calculate the mean deviation of the vector `LH`. 

```{r}
mean_deviation <- sum(abs(LH - arithmetic_mean)) / length(LH)
```

(b) There are two formulas for variance above. Use one to calculate the variance of vector `LH`, call this `var1`.

```{r}
var1 <- sum((LH - arithmetic_mean)^2) / (length(LH) - 1)
```

(c) Use the other variance formula to calculate the variance of vector `LH`, call this `var2`.

```{r}
var2 <- (sum((LH)^2) - ((sum((LH)))^2 / length(LH))) / (length(LH) - 1)
```

(d) Calculate the coefficient of variation.

```{r}
coef_var <- sqrt(var1) / arithmetic_mean
```

(e) Store the mean deviation, variance, and coefficient of variation of the vector `LH` in a vector called `LH_dispersion`. 

```{r}
LH_dispersion <- c(mean_deviation, var1, coef_var)
```

(f) Name each element in the vector `LH_dispersion` to clearly indicate which statistic is which. 

```{r}
names(LH_dispersion) <- c("mean deviation", "variation", "variation coeffient")
```

(g) Use an operator to check if `var1`  and `var2` are equal.  

```{r}
var1 == var2
```

(h) The two formulas for variance are identical mathematically.  Discuss why the two objects `var1` and `var2` might not have been equal.

The two objects of var1 and var2 may have not been equal because of disrepencies in the way both of the formulas are calculated as despite both formulas being mathematically exact, the values may round in certain ways which subtly changes the values between them while staying similar.
