---
title: "Homework #2 (40 pts)"
author: "Jaideep Bollu"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
 
## Instructions
- If you wish to use any additional packages beyond "Base R" (the default) then email Jericho (`jlaws011@ucr.edu`) for permission. 
- Use R to answer the questions below. 
- Check Canvas regularly for clarification on questions.  

# Questions 

**Q1: Arcade**

(a) Create the following objects in R:
+ Objects:
  - `color`: a factor of length 9 that has these elements in order: red, green, red, blue, red, blue, blue, red, green
  - `score`: a 3x3 matrix with the following elements in order by column: 14, 2, 5, 7, 10, 3, 7, 11, 8
  - `prizes`: a vector of three strings: "plush", "keychain", "pixy stix"
  
```{r}
color <- factor(c("red", "green", "red", "blue", "red", "blue", "blue", "red", "green"))

score <- matrix(c(14, 2, 5, 7, 10, 3, 7, 11, 8), nrow = 3, ncol = 3)

prizes <- c("plush", "keychain", "pixy stix")
```
  
(b) Get the column means of the matrix `score`.

```{r}
column_means <- colMeans(score)
column_means
```

(c) Create a list named *arcade* with the following elements, in order: `color`, `score`, and `prizes`. Make sure to name the elements in your list. Use the `str()` command to display the structure of the list.

```{r}
arcade <- list(color = "color", score = "score", prizes = "prizes")
str(arcade)
```

(d) Access and display the second element of the list using its index number. 

```{r}
arcade[[2]]
```

(e) Access and display the second element of the list using its name. 

```{r}
arcade$score
```

For the next part, please run this code before working on the problem.

```{r}
set.seed(13)
indices = sample(1:3, size = 9, replace = T)
```

(f) Use the `indices` object to subset the `prizes` object, which in turn will create a new object `prizeWon`. This object indicates what a person won from one three prizes in the `prize` object. 

```{r}
prizeWon <- prizes[indices]
prizeWon
```

(g) Create a DataFrame object named `teamPrize` which consists of the `color` object from part (a), a vectorized object of the `score` object from part (a), and the `prizeWon` object from part (f) as columns. Make sure to name the columns appropriately.

```{r}
score <- as.vector(score)
teamPrize <- data.frame(color, score, prizeWon)
names(teamPrize) <- c("color", "score", "prizeWon")
teamPrize
```

(h) In a method of your choice that has been discussed in class, find the mean and standard deviation of the scores for each color and output those results.

```{r}
red_score <- subset(teamPrize, teamPrize$color == "red")
mean_red <- mean(red_score$score)
sd_red <- sd(red_score$score)

blue_score <- subset(teamPrize, teamPrize$color == "blue")
mean_blue <- mean(blue_score$score)
sd_blue <- sd(blue_score$score)

green_score <- subset(teamPrize, teamPrize$color == "green")
mean_green <- mean(green_score$score)
sd_green <- sd(green_score$score)

mean_red
sd_red
mean_blue
sd_blue
mean_green
sd_green
```

**Q2: Indexing with Pokemon**

(a) Download the data set `Pokemon.csv` from *ELearn*.  This data set comes from [Kaggle.com](https://www.kaggle.com/abcsds/pokemon). Name this data set `pokemon`. Display the first 10 lines of this data set.

```{r}
pokemon <- read.csv("Pokemon.csv")
head(pokemon, 10)
```

(b) Use commands in R to determine how many Pokemon are from generation 2 that have an attack greater than 120.

```{r}
pokemon_gen2 <- subset(pokemon, pokemon$Generation == 2 & pokemon$Attack > 120)
pokemon_gen2_count <- nrow(pokemon_gen2)
pokemon_gen2_count
```

(c) Display the 17th largest value in the `HP` column of the original data set using R commands.

```{r}
ordered_hp <- pokemon$HP[order(pokemon$HP, decreasing = TRUE)]
pokemon_17 <- ordered_hp[17]
pokemon_17
```

(d) Create a new factor column in the Pokemon data set called `Mentality`. If the difference between the variable `Attack` is larger than the variable `Defense`, assign this character's `Mentality = Aggressive`.  If `Attack` is smaller than `Defense` then assign  `Mentality = Protective`.  If they are equal assign `Mentality = Balanced`. Display the first 5 rows of the Pokemon data set with this new variable. HINT: Create a new column and then using indexing and operators.

```{r}
pokemon$Mentality <- ifelse(pokemon$Attack > pokemon$Defense, "Aggressive", ifelse(pokemon$Attack < pokemon$Defense, "Protective", "Balanced"))

head(pokemon, 5)
```

**Q3: Data Manipulation with Pokemon**

(a) Create a new data set called `pokemon2` which only contains the columns:  `Name`, `Total`, `HP`, `Attack`,`Defense`, `Sp. Atk`, `Sp. Def`. Use the `head()` function to display the first few lines of this new data set.

```{r}

colnames(pokemon)

pokemon2 <- pokemon[c("Name", "Total", "HP", "Attack", "Defense", "Sp..Atk", "Sp..Def")]
head(pokemon2)
```

(b) Create a new data set called `pokemon3` which only contains the legendary Pokemon that have `Total` greater than the average (arithmetic mean) `Total` value.  In addition, have this data set only contain the following columns:  `Name`, `Type 1`, `Total`,  `Speed`,  `Generation`. Display rows 11-18 from this new data set.

```{r}
total_mean <- mean(pokemon$Total)
pokemon3 <- pokemon[pokemon$Legendary == "True" & pokemon$Total > total_mean, c("Name", "Type.1", "Total", "Speed", "Generation")]


pokemon3[11:18, ]
```

(c) The Pokemon data set contains `NA` values.  Create a new data set called `pokemon4` which only contains complete cases.  That is, this data set only contains the rows which have no `NA` values. Display the first 10 rows of this new data set.

```{r}
pokemon4 <- na.omit(pokemon)
head(pokemon4, 10)
```

(d) Determine the Pokemon characters have the 10 largest values in the `Total` column. Display *only the names* of these Pokemon.

```{r}
ordered_total <- pokemon$Total[order(pokemon$Total, decreasing = TRUE)]
large_total <- pokemon[ordered_total, c("Name")]
head(large_total, 10)
```

**Q4: Build Your Own Functions**

Some parts of this problem will refer to material from previous homework assignments. Please refer to the previous homework assignments to aid in your answers below. You will also use the `iris` data set in this question.

(a) Create a function that calculates and returns the harmonic mean for any arbitrary numeric vector, call it `harm.mean`.

```{r}

harm.mean <- function(x) {
  return((sum(1 / x) / length(x))^(-1))
}

xe <- c(10, 20, 30)

harm.mean(xe)

```

(b) Create a function that calculates and returns the geometric mean for any arbitrary numeric vector, call it `geom.mean`.

```{r}
geom.mean <- function(x) {
  return(exp(sum(log(x)) / length(x)))
}

geom.mean(xe)
```

(c) Create a function that calculates and returns the median for any arbitrary numeric vector, call it `my.median`. This function should only have one argument, a vector containing the values you wish to find the median of. DO NOT USE the `median()` function. HINT: The mod function will be useful here.

```{r}
my.median <- function(x) {
  sort_x <- sort(x)
  
  if(length(x) %% 2 == 1) {
    return(sort_x[((length(x) + 1) / 2)])
  }
  
  else {
    return(((sort_x[length(x) / 2]) + (sort_x[(length(x) / 2) + 1])) / 2) 
  }
  
}

x <- c(10, 20, 30, 40)
my.median(xe)
my.median(x)
```

(d) Use the `harm.mean()`, `geom.mean()`, and `my.median()` functions you created to calculate the harmonic mean, geometric mean, and median of the `Sepal.Width` variable in the `iris` dataset.  Store the results of these functions in a *named* vector and display the vector.

```{r}
data("iris")
head(iris, 10)

Width_results <- c(harm.mean(iris$Sepal.Width), geom.mean(iris$Sepal.Width), my.median(iris$Sepal.Width))
names(Width_results) <- c("Harmonic Mean", "Geometric Mean", "Median")
Width_results
```

(e) Using the `iris` data with `Species` equal to `versicolor`,  determine the median of the `Petal.Length` variable by using the function created in part (c).

```{r}
iris_versicolor <- subset(iris, Species == "versicolor")
Petal.Length_median <- my.median(iris_versicolor$Petal.Length)
Petal.Length_median
```
