---
title: "Discussion 2"
author: "Jaideep Bollu"
output:
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Suppose you have collected answers to the following survey questions from five respondents:

- What is your favorite nearby restaurant?
- What is your favorite subject?
- What is your favorite number?
- What year are you?
- Are you from California?

Here are the following results:

- Restaurant: Sushi Times, 7 Leaves, Manna Grill, 7 Leaves, Denny's
- Subject: English, Math, Math, Biology, Spanish
- Number: 7, 13, 0, 8, 5
- Year: First, Fourth, Third, Third, Second
- California?: Yes, Yes, No, Yes, No

Use the following responses to answer the questions below:

1) Create a logical vector named `cali` that consists of responses to the question about being from California. Let "yes" responses be represented as `TRUE`. Print out the result. (1 pt)
```{r}
cali <- c(TRUE, TRUE, FALSE, TRUE, FALSE)
print(cali)
```

2) Use the `as.double()` function to transform `cali` into a double vector called `dbl_cali`. Print out the result.  (1 pt)
```{r}
dbl_cali <- as.double(cali)
print(dbl_cali)
```

3) Comment on what you noticed about this conversion in the previous prompt. (1 pt)

I noticed that the TRUE values were converted into values of 1 when the vector was changed from a logical vector to a double vector and FALSE values were likewise converted to 0.

4) Create an ordered factor named `year` that stores the responses to the year question with the following hierarchy: first $\rightarrow$ second $\rightarrow$ third $\rightarrow$ fourth. Print out the result. (1 pt)
```{r}
yearVal <- c("First", "Fourth", "Third", "Third", "Second")
year <- factor(yearVal, ordered = TRUE, levels = c("First", "Second", "Third", "Fourth"))
print(year)
```

5) For the rest of the survey questions, please create appropriate vectors named `rest`, `subject`, and `number` respectively. (1 pt)
```{r}
rest <- c("Sushi Times", "7 Leaves", "Manna Grill", "7 Leaves", "Denny's")
subject <- c("English", "Math", "Math", "Biology", "Spanish")
number <- c("7", "13", "0", "8", "5")
```

6) Construct a data frame named `survey` that includes the `rest`, `subject`, `number`, `cali`, and `year` vectors. Print out the data frame. (1 pt)
```{r}
survey <- data.frame(rest, subject, number, cali, year)
print(survey)
```

7) Provide two ways for accessing the third response regarding `rest` from the `survey` data frame. (1.5 pts)
```{r}
survey[3, "rest"]

survey$rest[3]
```

8) **Extra Credit**: In one line of code, sort the data frame based on the `number` column from least to greatest. (0.5 pts)
```{r}
survey[order(as.numeric(survey$number)), ]
```




